---
title: "Example"
author: "Jean Steve TAMO"
date: "2024-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install and Usage
```{r}
library(PenFFR)
```


## Example

This is a basic example which shows you how to perform a linear function-on-function regression in different context :
* Non penalized concurrent model
* Penalized concurrent model
* Non penalized integral model
* Penalized integral model
* Non Penalized Mixture-of-Experts concurrent model
* Penalized Mixture-of-Experts concurrent model

The example is based on the Canadian weather dataset.
  
  
### Load the Canadian weather dataset
```{r}
data(fda::CanadianWeather)

raw.temp <- t(fda::CanadianWeather[["dailyAv"]][,,1])
raw.prec <- t(fda::CanadianWeather[["dailyAv"]][,,3])
raw.clim <- c(rep(1,15), rep(2,9), rep(3,5), rep(2,3), rep(4,3))
n <- 35 # number of individuals
m <- 365 # number of observations points
d <- 1 # number of predictors
obs.grid <- ((1:365)-1)/(365-1) # observation grid
```

### Functional expansion of Temperature (cubic splines with 148 nodes)
```{r}
temp.fd <- my_functrep(data = raw.temp,
                       nbasis = 150,
                       name.obs = fda::CanadianWeather[["place"]])
```
  
  
To perform the different models, the first observation of the dataset will be use as the test set.
  
  
### Load the PenFFR function
  
For the concurrent model, we used $15$ basis functions for the functional parameters in both penalized and non penalized models.
### Non penalized version of concurrent linear model
```{r}
fofreg1 <- PenFFR(Y.mat = raw.prec[-1,],
                  X_fd.list = list(temp.fd[-1]),
                  X.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[-1,],
                  nbasis = 15, pen = F,
                  model.type = "concurrent", K.mixture = "No")
```
  
  
### Penalized version of concurrent linear model
```{r}
fofreg2 <- PenFFR(Y.mat = raw.prec[-1,],
                  X_fd.list = list(temp.fd[-1]),
                  X.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[-1,],
                  nbasis = 15, pen = T,
                  model.type = "concurrent", K.mixture = "No")
```
  
  
For the integral versions, we uses less basis functions than concurrent models and we set to $8$.
### Non penalized version of integral model
```{r, cache=TRUE}
fofreg3 <- PenFFR(Y.mat = raw.prec[-1,],
                  X_fd.list = list(temp.fd[-1]),
                  X.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[-1,],
                  nbasis = 8, pen = F,
                  model.type = "integral", K.mixture = "No")
```
  
  
### Penalize version of integral model
```{r, cache=TRUE}
fofreg4 <- PenFFR(Y.mat = raw.prec[-1,],
                  X_fd.list = list(temp.fd[-1]),
                  X.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[-1,],
                  nbasis = 8, pen = T,
                  model.type = "integral", K.mixture = "No")
```
  
  
### Non Penalized concurrent Mixture-of-Experts model
```{r, cache=TRUE}
fofreg5 <- PenFFR(Y.mat = raw.prec[-1,],
                  X_fd.list = list(temp.fd[-1]),
                  X.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[-1,],
                  nbasis = 10, pen = F,
                  model.type = "concurrent", K.mixture = "BIC")
```
  
  
### Penalized concurrent Mixture-of-Experts model
```{r, cache=TRUE}
fofreg6 <- PenFFR(Y.mat = raw.prec[-1,],
                  X_fd.list = list(temp.fd[-1]),
                  X.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[-1,],
                  nbasis = 10, pen = T,
                  model.type = "concurrent", K.mixture = "BIC")
```

  
### Todo
Integral Mixture-of-Experts model coming soon !!!
  
  
  
## Visualization of parameters
Here we are going to give the graphical representations of the estimated functional parameters.
  
  
### Intercept of the different versions
```{r}
plot(fofreg1$beta.fd[[1]], xlab = "Days of Year", cex.lab = 1, ylim = c(0,3),
     ylab = latex2exp::TeX("$\\hat{$\\beta}_0(t)"), col = 1, lwd = 2,
     xaxt = "n")
axis(1, at = (c(1,32,60,91,121,152,182,213,244,274,305,335)-1)/364,
     labels = substr(month.name, 1, 3))
lines(fofreg2$beta.fd[[1]], col = "red", lwd = 2)
lines(fofreg3$beta.fd[[1]], col = "blue", lwd = 2)
lines(fofreg4$beta.fd[[1]], col = "green", lwd = 2)
abline(h = 0, lty = 3)
```
  
  

### Temperature effect parameters
```{r}
plot(fofreg1$beta.fd[[2]], xlab = "Days of Year", cex.lab = 1, ylim = c(-0.01,0.1),
     ylab = latex2exp::TeX("$\\hat{$\\beta}_1(t)"), col = 1, lwd = 2,
     xaxt = "n")
axis(1, at = (c(1,32,60,91,121,152,182,213,244,274,305,335)-1)/364,
     labels = substr(month.name, 1, 3))
lines(fofreg2$beta.fd[[2]], col = "red", lwd = 2)

tmp <- fda::eval.bifd(sevalarg = obs.grid,
                 tevalarg = obs.grid,
                 bifd = fofreg3$beta.fd[[2]])
persp(x = obs.grid, y = obs.grid, z = tmp, theta = 45, phi = -5,
      xlab = "s", ylab = "t", ticktype = "detailed",
      main = " ", nticks = 4, zlab = "", col = "blue")
tmp <- fda::eval.bifd(sevalarg = obs.grid,
                 tevalarg = obs.grid,
                 bifd = fofreg4$beta.fd[[2]])
persp(x = obs.grid, y = obs.grid, z = tmp, theta = 45, phi = 5,
      xlab = "s", ylab = "t", ticktype = "detailed",
      main = " ", nticks = 4, zlab = "", col = "green")

```
  
  
## Prediction
  
```{r}
pred1 <- pred.PenFFR(model = fofreg1,
                      newX_fd.list = list(temp.fd[1]),
                      newX.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[1,],
                     model.type = "concurrent", K.mixture = "No")
pred1 <- matrix(pred1, ncol = m, byrow = T)

pred2 <- pred.PenFFR(model = fofreg2,
                      newX_fd.list = list(temp.fd[1]),
                      newX.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[1,],
                     model.type = "concurrent", K.mixture = "No")
pred2 <- matrix(pred2, ncol = m, byrow = T)

pred3 <- pred.PenFFR(model = fofreg3,
                      newX_fd.list = list(temp.fd[1]),
                      newX.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[1,],
                     model.type = "integral", K.mixture = "No")
pred3 <- matrix(pred3, ncol = m, byrow = T)

pred4 <- pred.PenFFR(model = fofreg4,
                     newX_fd.list = list(temp.fd[1]),
                     newX.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[1,],
                     model.type = "integral", K.mixture = "No")
pred4 <- matrix(pred4, ncol = m, byrow = T)

pred5 <- pred.PenFFR(model = fofreg5,
                      newX_fd.list = list(temp.fd[1]),
                      newX.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[1,],
                     model.type = "concurrent", K.mixture = "BIC")
pred5 <- matrix(pred5, ncol = m, byrow = T)

pred6 <- pred.PenFFR(model = fofreg6,
                     newX_fd.list = list(temp.fd[1]),
                     newX.scal = data.frame(unlist(fda::CanadianWeather[["coordinates"]]))[1,],
                     model.type = "concurrent", K.mixture = "BIC")
pred6 <- matrix(pred6, ncol = m, byrow = T)


```
  
  
  
## Visualization of predictions
```{r}
plot(obs.grid, raw.prec[1,], xlab = "Days of Year", cex.lab = 1, ylim = c(-2,2),
     ylab = "log precipitations mm", type = "p",
     pch = 20, lwd=0.2, xaxt = "n")
axis(1, at = (c(1,32,60,91,121,152,182,213,244,274,305,335)-1)/364,
     labels = substr(month.name, 1, 3))
lines(obs.grid, pred1[1,], lwd = 2)
lines(obs.grid, pred2[1,], col = "red", lwd = 2)
lines(obs.grid, pred3[1,], col = "blue", lwd = 2)
lines(obs.grid, pred4[1,], col = "green", lwd = 2)
```

## Prediction error
We give here the predictive error of each model through the integrated square error given by : $$\mathrm{ISE}_i = \int_0^{365}\,\Big(\mathrm{Y}_i(t) - \widehat{\beta}_{(\text{-}i)}\,\mathrm{X}_i(t)\Big)^2\,dt$$
where the predictor $\mathrm{X}_i(.)$ derives from the noisy daily temperature measurements; the functional response $\mathrm{Y}_i(.)$ is the log daily precipitation and  $\widehat{\beta}_{(\text{-}i)}$ is the functional parameter estimated in the data set of all the observations except for the  $i^{th}$ observation.
  
For sake of reducing the computational burden, instead of the ISE, the $\mathrm{L}^2$-norm between the actual and prediction values on a grid of values $t$ is used as a surrogate. It is given by:
$$\begin{eqnarray}
    \widehat{\mathrm{ISE}}_i = \sum_{j=1}^{365}\Big(\mathrm{Y}_i(j) - \widehat{\beta}_{(\text{-}i)}\,\mathrm{X}_i(j)\Big)^2.
\end{eqnarray}$$
  
### Non penalized concurrent model
```{r}
err1 <- my_ISE(Y.act = raw.prec[1,],
               Y.hat = pred1)
err1
```
  
### Penalized concurrent model
```{r}
err2 <- my_ISE(Y.act = raw.prec[1,],
               Y.hat = pred2)
err2
```
  
### Non penalized integral model
```{r}
err3 <- my_ISE(Y.act = raw.prec[1,],
               Y.hat = pred3)
err3
```
  
### Penalized integral model
```{r}
err4 <- my_ISE(Y.act = raw.prec[1,],
               Y.hat = pred4)
err4
```

### Non penalized concurrent mixture-of-experts model
```{r}
err5 <- my_ISE(Y.act = raw.prec[1,],
               Y.hat = pred5)
err5
```
  
### Penalized concurrent mixture-of-experts model
```{r}
err6 <- my_ISE(Y.act = raw.prec[1,],
               Y.hat = pred6)
err6
```
